\sick, the spectroscopic inference crank

### 2014-08-24 document/manuscript.tex
Updated manuscript

 \begin{equation}
\mathcal{L} = \prod_{i=1}^{N}\,\left[\left(1 - P_{o}\right)\times{}p_{model}\left(F_i|\lambda_i,\sigma_{i},\bm{\kappa}\right) + P_{o}\times{}p_{outlier}\left(F_i|\lambda_i,\sigma_i,\bm{\kappa},V_{o},P_o\right)\right]
\end{equation}

@@

\begin{equation}
p_{outlier}\left(F_i|\lambda_i,\sigma_i,\bm{\kappa},V_{o},P_o\right) = \frac{1}{\sqrt{2\pi\left(s_{i}^2 + V_{o}^2\right)}} \exp\left(-\frac{[F_i - C_i]^2}{2\left[s_{i}^2 + V_{o}^2\right]}\right)
\end{equation}

@@

\begin{equation}
\mathcal{L} = \prod_{i=1}^{N} \left[ \frac{1-P_o}{\sqrt{2\pi{}s_{i}^2}}\,\exp\,\left(-\frac{[F_i - E_i]^2}{2s_{i}^{2}}\right) + \frac{P_o}{\sqrt{2\pi\left[s_{i}^2 + V_o\right]}}\,\exp\,\left(-\frac{[F_i - C_i]^2}{2\left[s_{i}^{2} + V_o\right]}\right)\right]
\label{eq:full_likelihood}
\end{equation}

I define the full parameter space with $\bm{\theta} \equiv \left(\bm{\psi},\{z,\sigma_s,\{c_{b_k}\}_{k=0}^{m},f\}_{b=0}^{N_{c}},V_o,P_o\right)$. From Bayes theorem the posterior probability 
distribution for $\bm{\theta}$ (up to a constant) is given by
\begin{eqnarray}
\mathcal{P} & \propto & likelihood \times prior \nonumber \\
p(\bm{\theta}|\{F_i\}_{i=1}^{N}) & \propto & p(\{F_i\}_{i=1}^{N}|\bm{\theta})\,\times\,p(\bm{\theta})
\label{eq:probability}
\end{eqnarray}

@@

The random scattering and numerical optimisation steps efficiently provide an 
accurate estimate of the optimal parameters $\bm{\theta_{opt}}$. Once the optimisation
 step is complete, \sick{} employs the affine-invariant ensemble sampler proposed 
 by \citet{goodman;weare}, and implemented by \citet{emcee}. The Metropolis-Hastings 
 MCMC algorithm is employed by default. \sick{} allows for the model settings to 
 be specified in a human-readable \textsc{yaml}- or \textsc{json}-formatted 
 configuration file\footnote{The reader is referred to the online documentation 
 for an example.}, where the number of Goodman \& Weare walkers can be specified, 
 as well as the number of samples to perform. When the optimisation step is used, 
 the initial points are taken from a small multi-dimensional ball around the 
 optimised parameters $\bm{\theta_{opt}}$. If the optimisation step is not 
 performed, initial values for the MCMC walkers are drawn in the same way as 
 described for the scattering process.

@@

\begin{eqnarray}
p\left(\bm{\psi}_{dim}\right) &=& \mathcal{U}\left(\min\left[\bm{\psi}_{dim}\right], \max\left[\bm{\psi}_{dim}\right]\right) \\
p\left(P_o\right) &=& \mathcal{U}\left(0, 1\right) \\
p\left(\log{f}\right) &=& \mathcal{U}\left(-10, 1\right) \\
p\left(V_o,\sigma_s\right) &=& \left\{
\begin{array}{c l}      
    1\,, &\mbox{for values greater than zero}\\
    0\,, &\mbox{otherwise}
\end{array}\right. \\
p\left(z,\{c_k\}_{k=0}^{m}\right) &=& 1 \\
\label{eq:default_priors}
\end{eqnarray}

@@

\begin{figure*}
\label{fig:chains}
\includegraphics[height=\textheight]{chains.pdf}
\caption{Points sampled by the 200 walkers at each step during the self-consistent 
inference test. The true values are marked in blue. The first 1250 steps are 
discarded as the burn-in period.}
\end{figure*}

\begin{figure*}
\label{fig:corner-inference}
\includegraphics[width=\textwidth,height=\textwidth]{corner.pdf}
\caption{Marginalised posterior distributions for all parameters $\bm{\theta}$ 
from a faux observation with spectral resolution $\mathcal{R} \sim 10,000$ and 
S/N ratio $\sim{}7$\,pixel$^{-1}$. True values are marked in blue. This figure 
demonstrates that precise inference of stellar parameters can be made with 
high-resolution spectra, even in the presence of substantial noise.}
\end{figure*}

@@

\begin{figure*}
\label{fig:spectrum-inference}
\includegraphics[width=\textwidth]{spectrum.pdf}
\caption{A faux observed spectrum (black) of $\mathcal{R} \sim 10000$ and S/N 
ratio of $\sim7$ with an underestimated variance by $\sim$10\% which was used for 
the self-consistent inference test. The recovered maximum likelihood model spectrum 
is shown in red. }
\end{figure*}

@@

\begin{figure*}
\label{fig:solar}
\includegraphics[width=\textwidth,height=\textwidth]{solar.pdf}
\caption{Stellar parameter ($\bm{\psi}$: $T_{\rm eff}$, $\log{g}$, [Fe/H], 
[$\alpha$/Fe])  posterior distributions for the GIRAFFE/FLAMES twilight spectrum. 
Nuisance parameters are not shown. Maximum likelihood values for each parameter 
are drawn in red, and the 16\%, 50\% and 84\% quantiles are shown as dashed lines. 
The posterior distributions are precise, and close to the accepted values (not 
shown) of $T_{\rm eff} = 5777$\,K, $\log{g} = 4.445$, slightly discrepant from 
[$\alpha$/Fe] = 0, and significantly deviant from [Fe/H] = 0. This test 
illustrates the precision achievable using \sick, but highlights the need for 
accurate models.}
\end{figure*}

@@

\begin{figure}[h!]
\label{fig:clusters}
\includegraphics[width=0.5\textwidth]{clusters.pdf}
\caption{Metallicities of cluster stars inferred from noisy, low-resolution 
spectra obtained with the AAOmega spectrograph. The agreement with the 
literature values (see text) is very reasonable.}
\end{figure}

@@

With applications to both high- and low-resolution stellar spectra, I have 
demonstrated that the simultaneous incorporation of these effects leads to 
remarkable improvements in both accuracy and precision. While the examples 
presented here have focussed on stellar spectra, the code is ambivalent about 
\textit{what} the astrophysical parameters describe: the framework can be easily 
used for any kind of quantifiable astrophysical process. The code is MIT-licensed 
(freely distributable, open source) and has an extensive automatic testing suite, 
which includes regular reproduction of the examples presented in this \article{}. 
Complete documentation is available online\footnote{astrowizici.st/sick}, 
which includes a number of additional examples and tutorials. In the spirit of 
full scientific reproducibility, the online documentation is complemented with the 
files necessary to reproduce all of the examples presented in this \article{}. 
The source code is distributed using \textsc{git}, and hosted online at 
GitHub\footnote{github.com/andycasey/sick}.

>>>>

 \begin{multline}
\mathcal{L} = \prod_{i=1}^{N}\,\left[\left(1 - P_{o}\right)\times{}p_{model}\left(F_i|\lambda_i,\sigma_{i},\bm{\kappa}\right)\right. \\
+ \left. P_{o}\times{}p_{outlier}\left(F_i|\lambda_i,\sigma_i,\bm{\kappa},V_{o},P_o\right)\right]
\end{multline}

@@

\begin{multline}
p_{outlier}\left(F_i|\lambda_i,\sigma_i,\bm{\kappa},V_{o},P_o\right) = \dots \\
  \frac{1}{\sqrt{2\pi\left(s_{i}^2 + V_{o}^2\right)}} \exp\left(-\frac{[F_i - C_i]^2}{2\left[s_{i}^2 + V_{o}^2\right]}\right)
\end{multline}

@@

\begin{multline}
\mathcal{L} = \prod_{i=1}^{N} \left[ \frac{1-P_o}{\sqrt{2\pi{}s_{i}^2}}\exp\left(-\frac{[F_i - E_i]^2}{2s_{i}^{2}}\right) \right.\\
\left. + \frac{P_o}{\sqrt{2\pi\left[s_{i}^2 + V_o\right]}}\exp\left(-\frac{[F_i - C_i]^2}{2\left[s_{i}^{2} + V_o\right]}\right)\right]
\label{eq:full_likelihood}
\end{multline}

I define the full parameter space with ${\bm{\theta} \equiv \left(\bm{\psi},\{z,\sigma_s,\{c_{b_k}\}_{k=0}^{m},f\}_{b=0}^{N_{c}},V_o,P_o\right)}$. From Bayes theorem the posterior probability 
distribution for $\bm{\theta}$ (up to a constant) is given by

\begin{eqnarray}
\mathcal{P} & \propto & likelihood \times prior \nonumber \\
p(\bm{\theta}|\{F_i\}_{i=1}^{N}) & \propto & p(\{F_i\}_{i=1}^{N}|\bm{\theta})\,\times\,p(\bm{\theta})
\label{eq:probability}
\end{eqnarray}

@@

The random scattering and numerical optimisation steps efficiently provide an 
accurate estimate of the optimal parameters $\bm{\theta_{opt}}$. Once the optimisation
 step is complete, \sick{} employs the affine-invariant ensemble sampler proposed 
 by \citet{goodman;weare}, and implemented by \citet{emcee}. The Metropolis-Hastings 
 MCMC algorithm is employed by default. \sick{} allows for the model settings to 
 be specified in a human-readable \textsc{yaml}- or \textsc{json}-formatted 
 configuration file\footnote{The reader is referred to the online documentation at http://astrowizici.st/sick/ 
 for an example.}, where the number of Goodman \& Weare walkers can be specified, 
 as well as the number of samples to perform. When the optimisation step is used, 
 the initial points are taken from a small multi-dimensional ball around the 
 optimised parameters $\bm{\theta_{opt}}$. If the optimisation step is not 
 performed, initial values for the MCMC walkers are drawn in the same way as 
 described for the scattering process.

@@

\begin{eqnarray}
p\left(\bm{\psi}_{dim}\right) &\,=\,& \mathcal{U}\left(\min\left[\bm{\psi}_{dim}\right], \max\left[\bm{\psi}_{dim}\right]\right) \\
p\left(P_o\right) &\,=\,& \mathcal{U}\left(0, 1\right) \\
p\left(\log{f}\right) &\,=\,& \mathcal{U}\left(-10, 1\right) \\
p\left(V_o,\sigma_s\right) &\,=\,& \left\{
\begin{array}{c l}      
    1\,, &\mbox{for values greater than zero}\\
    0\,, &\mbox{otherwise}
\end{array}\right. \\
p\left(z,\{c_k\}_{k=0}^{m}\right) &\,=\,& 1 \\
\label{eq:default_priors}
\end{eqnarray}

@@

\begin{figure*}
\label{fig:chains}
\includegraphics[height=\textheight]{figures/chains.pdf}
\caption{Points sampled by the 200 walkers at each step during the self-consistent 
inference test. The true values are marked in blue. The first 1250 steps are 
discarded as the burn-in period.}
\end{figure*}

\begin{figure*}
\label{fig:corner-inference}
\includegraphics[width=\textwidth,height=\textwidth]{figures/corner.pdf}
\caption{Marginalised posterior distributions for all parameters $\bm{\theta}$ 
from a faux observation with spectral resolution $\mathcal{R} \sim 10,000$ and 
S/N ratio $\sim{}7$\,pixel$^{-1}$. True values are marked in blue. This figure 
demonstrates that precise inference of stellar parameters can be made with 
high-resolution spectra, even in the presence of substantial noise.}
\end{figure*}

@@

\begin{figure*}
\label{fig:spectrum-inference}
\includegraphics[width=\textwidth]{figures/spectrum.pdf}
\caption{A faux observed spectrum (black) of $\mathcal{R} \sim 10000$ and S/N 
ratio of $\sim7$ with an underestimated variance by $\sim$10\% which was used for 
the self-consistent inference test. The recovered maximum likelihood model spectrum 
is shown in red. }
\end{figure*}

@@

\begin{figure*}
\label{fig:solar}
\includegraphics[width=\textwidth,height=\textwidth]{figures/solar.pdf}
\caption{Stellar parameter ($\bm{\psi}$: $T_{\rm eff}$, $\log{g}$, [Fe/H], 
[$\alpha$/Fe])  posterior distributions for the GIRAFFE/FLAMES twilight spectrum. 
Nuisance parameters are not shown. Maximum likelihood values for each parameter 
are drawn in red, and the 16\%, 50\% and 84\% quantiles are shown as dashed lines. 
The posterior distributions are precise, and close to the accepted values (not 
shown) of $T_{\rm eff} = 5777$\,K, $\log{g} = 4.445$, slightly discrepant from 
[$\alpha$/Fe] = 0, and significantly deviant from [Fe/H] = 0. This test 
illustrates the precision achievable using \sick, but highlights the need for 
accurate models.}
\end{figure*}

@@

\begin{figure}[h!]
\label{fig:clusters}
\includegraphics[width=0.5\textwidth]{figures/clusters.pdf}
\caption{Metallicities of cluster stars inferred from noisy, low-resolution 
spectra obtained with the AAOmega spectrograph. The agreement with the 
literature values (see text) is very reasonable.}
\end{figure}

@@

With applications to both high- and low-resolution stellar spectra, I have 
demonstrated that the simultaneous incorporation of these effects leads to 
remarkable improvements in both accuracy and precision. While the examples 
presented here have focussed on stellar spectra, the code is ambivalent about 
\textit{what} the astrophysical parameters describe: the framework can be easily 
used for any kind of quantifiable astrophysical process. The code is MIT-licensed 
(freely distributable, open source) and has an extensive automatic testing suite, 
which includes regular reproduction of the examples presented in this \article{}. 
Complete documentation is available online\footnote{astrowizici.st/sick}, 
which includes a number of additional examples and tutorials. In the spirit of 
scientific reproducibility, the online documentation is complemented with the 
files necessary to reproduce examples presented in this \article{}. 
The source code is distributed using \textsc{git}, and hosted online at 
GitHub\footnote{github.com/andycasey/sick}.
